Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-10-12 doc/JFLA2022/paper.tex
pass over abstract

\begin{abstract}
  \Hydras is a collaborative library on discrete math, written for the \coq  proof assistant, and developed under the \community project. The proof scripts are
  accompanied with an electronic book, made with the help of the \alectr litterate proving tool.
  We present the evolution of the mathematical contents since
  former presentations at JFLA meetings.
  Then, we show how the structure of the project is determined   by two  requirements which must be continuously satisfied.
   First, the Coq code needs to be maintained for compatibility with its dependencies (the Coq proof assistant and several Coq packages from inside and outside \community) and reverse dependencies (\community projects that depend on it).
   Second, the book needs to be maintained consistent with respect to the evolutions of the Coq code.
   %\TODO[Pierre]{Find a better/simpler sentence about the project's structure?}
\end{abstract}

@@

Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}. However, after a key mathematical definition or result is added to a library, it needs be \emph{documented} and \emph{maintained}.

Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions. Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB}. The former tends to be terse and dense and serves experienced users, while the latter is more long-winded and exhaustive and serves beginners. As a library changes and expands, all its documentation needs to be made consistent and complete. Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools~\cite{vanDoorn2020}, to pretty-print and check source code snippets and generate proof assistant output.

While books that document proof assistant libraries are valuable to beginners, the purpose of most current books is mainly instrumental, i.e., to teach a certain topic or technique. We believe that such books can become \emph{ends-in-themselves}---not just sources of learning, but also of inspiration and entertainment.

>>>>

\begin{abstract}
  \Hydras is a collaborative library of discrete mathematics,
  written for the \coq proof assistant and developed as part
  of the \community project on GitHub. The Coq code is accompanied
  by an electronic book, generated with the help of the \alectr
  literate proving tool. We present the evolution of the mathematical
  contents of the library since former presentations at JFLA meetings.
  Then, we describe how the structure of the project is determined
  by two requirements which must be continuously satisfied.
  First, the Coq code needs to be compatible with its ever-evolving
  dependencies (the Coq proof assistant and several Coq packages both from
  inside and outside \community) and its reverse dependencies (\community
  projects that depend on it). Second, the book needs to be consistent
  with the library Coq code, which undergoes frequent changes to improve
  structure and include new definitions and results.
\end{abstract}

@@

Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}. However, after a key mathematical definition or result is added to a library, it needs be \emph{documented} and \emph{maintained}~\cite{vanDoorn2020}.

Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions. Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB}. The former tends to be terse and dense and serves experienced users, while the latter is more long-winded and exhaustive and serves beginners. As a library changes and expands, all its documentation needs to be made consistent and complete. Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.

While books that document proof assistant libraries are valuable to beginners, the purpose of current books is mainly instrumental, i.e., to teach a certain topic or technique. We believe that such books can become \emph{ends in themselves}---not just sources of learning, but also sources of inspiration and entertainment.
